A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-15 manuscript.Rmd
add pagenumber for direct quote closes #138

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017]. <!--AB: direct quotes need a line number! See APA style guide-->
Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
First, there is a community that maintains pre-made images for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
Second, instead of saving a binary image of the software environment, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb) and sufficiently describes how to recreate a given container image.
However, Dockerfiles rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to keep a full image in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

@@

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019:

> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc. This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses. [...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses. [...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al. 2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998). Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
<!--AB you need to add the line numbers to the citation-->

>>>>

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174]. <!--APA requires no line number, but pagenumber or section heading-->

Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
First, there is a community that maintains pre-made images for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
Second, instead of saving a binary image of the software environment, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb) and sufficiently describes how to recreate a given container image.
However, Dockerfiles rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to keep a full image in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

@@

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019, p. 2:

> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc. This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses. [...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses. [...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al. 2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998). Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
<!--APA requires no line number, but pagenumber or section heading-->
